---
title: "ARIMA Analysis of Kalshi Prediction Market Prices"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(forecast)
library(tseries)
```

---

## Load and Prepare Data

```{r load-data}
trades <- read_csv("trades_clean.csv")

# parse timestamps, sort by contract then time
trades <- trades %>%
  mutate(created_time = as.POSIXct(created_time, tz = "UTC")) %>%
  arrange(ticker, created_time)

glimpse(trades)
```

---

## Pick a Representative Contract

Select a contract with enough trades to fit a time series model, use `yes_price` as the price series.

```{r select-contract}
# find contracts with the most trades
trade_counts <- trades %>%
  count(ticker, sort = TRUE)

print(head(trade_counts, 10))

# use the most-traded contract
selected_ticker <- trade_counts$ticker[1]
cat("Selected ticker:", selected_ticker, "\n")

contract_trades <- trades %>%
  filter(ticker == selected_ticker) %>%
  select(created_time, yes_price, outcome) %>%
  arrange(created_time)

cat("Number of trades:", nrow(contract_trades), "\n")
cat("Outcome:", unique(contract_trades$outcome), "\n")
```

---

## Price Series Plot

```{r plot-price}
ggplot(contract_trades, aes(x = created_time, y = yes_price)) +
  geom_line(color = "steelblue") +
  geom_point(size = 0.8, alpha = 0.5) +
  labs(
    title = paste("Yes Price Over Time:", selected_ticker),
    x = "Time",
    y = "Yes Price (implied probability)"
  ) +
  theme_minimal()
```

---

## ACF and PACF of Raw Price Series

Before fitting ARIMA, inspect autocorrelation in the raw price series.

```{r acf-raw}
price_ts <- ts(contract_trades$yes_price)

par(mfrow = c(1, 2))
acf(price_ts, main = "ACF of Yes Price", lag.max = 30)
pacf(price_ts, main = "PACF of Yes Price", lag.max = 30)
par(mfrow = c(1, 1))
```

---

## Stationarity Test (ADF)

An Augmented Dickey-Fuller test checks for a unit root, a hallmark of a random walk.

- **Null hypothesis**: series has a unit root (random walk / non-stationary)
- **p < 0.05**: reject the null means series is stationary

```{r adf-test}
adf_result <- adf.test(price_ts)
print(adf_result)
```

---

## Fit ARIMA Model

Select the best ARIMA(p, d, q) model based on AIC. 
A model with `d = 1` indicates the series needed differencing, which is consistent with a random walk.

```{r fit-arima}
arima_model <- auto.arima(price_ts, stepwise = FALSE, approximation = FALSE)
summary(arima_model)
```

---

## Residual Diagnostics

If prices are a random walk, ARIMA residuals should be white noise with no autocorrelation.
If residuals show autocorrelation, past prices carry predictive information.

```{r residual-diagnostics}
checkresiduals(arima_model)
```

```{r acf-residuals}
acf(residuals(arima_model),
    main = paste("ACF of ARIMA Residuals\n(", selected_ticker, ")"),
    lag.max = 30)
```

---

## Ljung-Box Test on Residuals

Formally tests whether residuals are white noise across multiple lags.

- **Null hypothesis**: residuals are white noise (random walk behavior)
- **p < 0.05**: significant autocorrelation remains means past prices are predictive

```{r ljung-box}
lb_test <- Box.test(residuals(arima_model), lag = 20, type = "Ljung-Box")
print(lb_test)
```

---

## Interpretation

```{r interpretation}
d_order <- arima_model$arma[6]  # differencing order from auto.arima
lb_p <- lb_test$p.value
adf_p <- adf_result$p.value

cat("ARIMA model order: ARIMA(",
    arima_model$arma[1], ",",
    arima_model$arma[6], ",",
    arima_model$arma[2], ")\n")
cat("ADF p-value:", round(adf_p, 4), "\n")
cat("Ljung-Box p-value:", round(lb_p, 4), "\n\n")
```

Conclusion: Significant autocorrelation remains in residuals (Ljung-Box p = 0).
Past prices carry predictive information, meaning the series is NOT a pure random walk.
This suggests price movements are autocorrelated, potentially reflecting
the market updating on new information in a non-random pattern.

---

## Repeat Across Multiple Contracts

Run the same analysis across several contracts to see whether the random walk result generalizes.

```{r multi-contract, fig.height=8}
# pick top 5 most-traded contracts
top_tickers <- trade_counts$ticker[1:5]

results <- map_dfr(top_tickers, function(tkr) {
  df <- trades %>%
    filter(ticker == tkr) %>%
    arrange(created_time)

  if (nrow(df) < 20) return(NULL)

  ts_data <- ts(df$yes_price)

  tryCatch({
    model <- auto.arima(ts_data, stepwise = TRUE, approximation = TRUE)
    lb <- Box.test(residuals(model), lag = 10, type = "Ljung-Box")
    adf <- adf.test(ts_data)

    tibble(
      ticker = tkr,
      n_trades = nrow(df),
      outcome = unique(df$outcome)[1],
      arima_order = paste0("(", model$arma[1], ",", model$arma[6], ",", model$arma[2], ")"),
      d_order = model$arma[6],
      adf_p = round(adf$p.value, 4),
      ljung_box_p = round(lb$p.value, 4),
      random_walk = ifelse(model$arma[6] >= 1 & lb$p.value > 0.05, "Yes", "No")
    )
  }, error = function(e) NULL)
})

print(results)
```

```{r acf-grid, fig.height=10}
# ACF plots for top contracts
par(mfrow = c(ceiling(length(top_tickers) / 2), 2))

for (tkr in top_tickers) {
  df <- trades %>%
    filter(ticker == tkr) %>%
    arrange(created_time)

  if (nrow(df) < 20) next

  tryCatch({
    model <- auto.arima(ts(df$yes_price), stepwise = TRUE, approximation = TRUE)
    acf(residuals(model),
        main = paste0(substr(tkr, 1, 30), "\nARIMA", paste0("(", model$arma[1], ",", model$arma[6], ",", model$arma[2], ")")),
        lag.max = 20)
  }, error = function(e) NULL)
}

par(mfrow = c(1, 1))
```
